---
title: "SDS315HW10"
author: "Max Allen"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(MatchIt)
library(mosaic)
library(moderndive)
library(effectsize)
library(gridExtra)
```

# Problem 1


I want to figure out whether there is an association between minority percentage and access to private insurance policies to prove existence of redlining. If there is an association, then I want to figure out exactly how big it might be and to what extent it matters.

For my approach, I am running a multivariate regression model and then bootstrapping it to get a 95% confidence interval of the effect of the minority percentage on FAIR policies in the ZIP code. The regression model will account for the amount of old houses in the ZIP code, the median income in the ZIP code, and the amount of fires per capita in the ZIP code. This allows me to see the individual effect of minority amount on policies, but also the effect that these other valuable variables play in access to private insurance for a region.

```{r echo=FALSE, message=FALSE}
redline = read_csv("redlining.csv")
redline %>% 
  ggplot(aes(x=minority, y=policies)) + geom_point() + geom_smooth(method="lm")

redline = redline %>%
  mutate(income_quartile = cut_number(income, 4))
redline = redline %>%
  mutate(age_quartile = cut_number(age, 4))
redline = redline %>%
  mutate(fire_quartile = cut_number(fire, 4))

plot1 = redline %>% 
  ggplot(aes(x=minority, y=policies, col=income_quartile)) + geom_point() + geom_smooth(method="lm", se=FALSE)
plot2 = redline %>% 
  ggplot(aes(x=minority, y=policies, col=age_quartile)) + geom_point() + geom_smooth(method="lm", se=FALSE)
plot3 = redline %>% 
  ggplot(aes(x=minority, y=policies, col=fire_quartile)) + geom_point() + geom_smooth(method="lm", se=FALSE)
combined_plot = grid.arrange(plot1, plot2, plot3)


boot_fit = do(1000)*lm(policies ~ minority + age + fire + income, data=resample(redline))
confint(boot_fit)
minority_conf = confint(boot_fit, level=.95)[2, 2:3]
```

Here I have a graph of the amount of FAIR policies for a ZIP code put against the minority make up of that area. It looks like a very strong, positive effect initially. However, after that I group these ZIP codes by income, age, and fire tendencies to show how these values greatly confound my variables of interest.
The confidence interval I found for the effect of minority on policies is (`r minority_conf`). While this seems like a very small increment, in the space of our variables of interest it is quite significant. Say, if there was a 50% increase in minority percentage, policies would go up by 50 times `r round(sum(minority_conf)/2, 4)`, or `r round(50*sum(minority_conf)/2, 4)` which is a meaningful increase in lack of access to private insurance. One might imagine the connection to be greater, but nonetheless, it is still apparent that redlining plays some roll in lack of access to private insurance policies.


# Problem 2

### Part A
```{r echo=FALSE, message = FALSE}
grocery = read_csv("groceries.csv")

grocery %>% 
  group_by(Store) %>% 
  summarize(Store_Mean_Price = mean(Price)) %>% 
  ggplot() + geom_col(aes(x=Store, y=Store_Mean_Price), fill='steelblue2') + coord_flip() + labs(y = "Mean Price of All Products ($)")

```

> This plot demonstrates average price of products across our grocery stores. They appear to differ a lot from store to store ranging from about $2 to $4 per product.

### Part B
```{r echo=FALSE}
grocery %>% 
  group_by(Product) %>% 
  summarize(Product_Count = n()) %>% 
  ggplot() + geom_col(aes(x=Product, y=Product_Count), fill="violet") + coord_flip() + labs(y="Amount of Stores Selling This Product")
```

> This graph is for the availability of products at different stores. As expected, milk and eggs are available everywhere while sugary items like mainstream ice creams and cereals are only in 4 stores.

### Part C
```{r echo=FALSE}
boot_groc = do(1000)*lm(Price ~ Product + Type, data=resample(grocery))
Grocer_vs_Convenience = confint(na.omit(boot_groc))[41, 2:3]
confint(na.omit(boot_groc))[41:44, 1:3]
```


Compared with ordinary grocery stores, convenience stores typically charge between `r round(Grocer_vs_Convenience, 2)[2]*-1` and `r round(Grocer_vs_Convenience, 2)[1]*-1` dollars more per product.

### Part D
```{r echo=FALSE}
boot_groc2 = do(1000)*lm(Price ~ Product + Store, data=resample(grocery))
groc_conf = confint(na.omit(boot_groc2))[41:53,]
ggplot(groc_conf) + geom_col(aes(x=name, y=estimate)) + coord_flip()
```

> This here is a graph of the estimated price increase/decrease per store using a 12 pack of Coke at  Albertson's as a baseline.

The two stores with the lowest price knock-up for the same products are Walmart and Kroger Fresh Fare. The two stores with the highest estimated price increase are Whole Foods and Wheatsville Food Coop.

### Part E
```{r echo=FALSE}
groc_conf[c(1,5),]
```

You can see above that the difference in price per product for HEB and Central Market is only about 10 cents a product. Other differences between stores reach up to a full dollar, so I would say that this is not a super meaningful price difference as one might expect. Because the confidence intervals are almost entirely overlapped, I conclude that HEB and Central Market charge similar prices for the same products.

### Part F
```{r echo=FALSE}
grocery = grocery %>% 
  mutate(Income10K = Income/10000)

boot_groc3 = do(1000)*lm(Price ~ Product + Income10K, data=resample(grocery))

confint(na.omit(boot_groc3))[40, ]

standardize_parameters(lm(Price ~ Product + Income10K, data=grocery))
```

The income in 10K coefficient has a negative sign which would indicate that poorer ZIP codes pay _more_ for each product on average. I know this because a negative slope would mean that price goes down as income goes up.
Using standardized coefficients, I found that a 1 standard deviation increase in the income of a ZIP code equates to a .03 standard deviation decrease in price that consumers in that ZIP code expect to pay while holding products constant.