---
title: "SDS315HW3"
author: "Max Allen"
date: "2024-02-01"
output: html_document
---
Max Allen
mca2773
Github -> https://github.com/mallen17/SDS315HW3.git

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)
library(kableExtra)
creatine = read_csv("creatinine.csv")
```

### Part A

```{r echo=FALSE}
exp_creat = lm(creatclear ~ age, data = creatine)
exp_cr = function(age){
  147.81 - .62 * age
}

```

For a fifty-five year old, we should expect a creatine clearance rate of `r exp_cr(55)`. I got this from forming a linear regression model between the two variables which gave me the expected relationship between age and creatine clearance.

### Part B

Creatine clearance decreases by .62mL/min for each year. I found this by finding the linear regression model and the coefficient of age.

### Part C

A 40 year old with a rate of 135 would have a residual of `r 135 - exp_cr(40)` while a 60 year old with a rate of 115 would have a residual of `r 115 - exp_cr(40)`. The 60 year old would be healthier because their creatine rate is `r (135 - exp_cr(60))`mL/min higher than should be expected for her age.


# Problem 2

### Introduction
The beta of a stock refers to its attachment to the success of the overall market. A beta of more than one indicates that the stock will react a lot more to the changes in the market, if the overall market booms then the stock will boom even more. However, this stands true for when the market busts. A beta of one would signify the average level of connection, naturally dependent on the market. A beta of less than one means that the stock is a lot less susceptible to market inflation. Take something such as Walmart, no matter what the market is like consumers will need their groceries to live which gives this stock somewhat of a shield against inflation. The beta would be the coefficient of our regression model, how much the stock goes up for a 100% return.

### Table
```{r echo=FALSE, warning=FALSE, message=FALSE}
market = read_csv("marketmodel.csv")
market = select(market, -Date)
spy = market$SPY
market = select(market, -SPY)
stock_tix = c("AAPL", "GOOG", "MRK", "JNJ", "WMT", "TGT")
regress = function(stock){
  lm(stock ~ spy)
}
determination = function(stock){
  cor(stock, spy) ^ 2
}

int_tib = c(1:6)
coef_tib = c(1:6)
determ_tib = c(1:6)

for (i in 1:6) {
  int_tib[i] = coef(regress(market[[stock_tix[i]]]))[1]
  coef_tib[i] = coef(regress(market[[stock_tix[i]]]))[2]
  determ_tib[i] = determination(market[[stock_tix[i]]])
}

clean_market = tibble(Ticker = colnames(market), Intercept = int_tib, Coefficient = coef_tib, RSquared = determ_tib)
clean_market %>% 
  kbl(format = "html", align = "c", label = "Popular Stocks Regressed Against The Market") %>% 
  kable_styling(bootstrap_options = "striped")

```

> The table here gives the linear regression models for each of these six stocks when compared against the market as a whole (we used the S&P 500 stock to track this). The coefficient column should be the beta of each stock while the rsquared column is the coefficient of determination, the percentage of y that can be explained by x.

### Conclusion

The lowest systematic risk of my six stocks is unsurprisingly Walmart while the highest is Apple, though it only barely has a beta over one. This all seems quite reasonable is Walmart is needed irrespective of inflation status and Apple would be hit a lot harder due to their high-priced tech products.


# Problem 3

```{r echo=FALSE, message=FALSE, warning=FALSE}
covid = read_csv("covid.csv")
italy = filter(covid, country=="Italy")
spain = filter(covid, country=="Spain")
covid %>% 
  ggplot() + geom_line(aes(x=days_since_first_death, y=deaths, col=country)) + labs(title="Covid Deaths Since Breakout", x="days after first covid death")


gr_sp = round(coef(lm(log(deaths) ~ days_since_first_death, data = spain))[2], 3)
dbl_time_sp = round(70/(gr_sp*100))
gr_it = round(coef(lm(log(deaths) ~ days_since_first_death, data = italy))[2], 3)
dbl_time_it = round(70/(gr_it*100))

```

This data displays the exponential relationship of covid deaths after the first covid fatality in Italy and Spain. In Spain, the estimated growth rate of deaths is `r gr_sp` where the doubling time is `r dbl_time_sp` days. In Italy, the estimated growth rate is `r gr_it` and the doubling time is `r dbl_time_it` days.


# Problem 4

```{r echo=FALSE, message=FALSE}
milk = read_csv("milk.csv")

milksticity = coef(lm(log(sales) ~ log(price), data = milk))[2]
```

The elasticity of milk demand is `r round(milksticity, 2)` which means that for 1% of price increase, milk sales go down `r round(-(milksticity), 2)`%. This makes sense as people would naturally buy a good deal less milk if it was too expensive and outside of their budget, however, it is still essential and always bought a decent amount. All I did was found the linear model between the log of milk sales and the log of milk prices, the coefficient yields elasticity.